---
title: "Lecture 9 - The Good, the Bad, the Ugly Notebook"
author: "Lindsay Waldrop"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

```{r setup}
library(ggplot2)
```

# Lecture 9 Notebook

## Assignment

Use the diamonds data set (in ggplot2 package) to answer the following questions (Wilke chapters associated with question). (More on the diamonds data set here: https://rpubs.com/GinaMoreno__/course1_4)

 1. Are flawless diamonds on average smaller than those with inclusions? (Ch 6)
 2. What is the distribution of price per carat of the diamonds in this set? (Ch 7)
 3. What features determine a diamond’s cost? (Ch 9)
 4. What proportion of diamonds in the set are of each cut? Does this change with clarity? (Ch 10, 11)
 5. How does depth relate to the price of a diamond? (Ch 12)
 
```{r}
data("diamonds")
```


Use the beavers data set (base R) to answer the following questions:

 6. Does the body temperature of a beaver vary with time? (Ch 13, 14)
 7. How much variation is associated with beaver temperature measurements?   
      (Ch 16)

```{r}
data("beavers")
# Note that the beavers data set contains two objects: beaver1 and beaver2

```
 
You will be assigned one question number to complete. 

Prepare a graph that best answers the question and post it to the discussion board. Include an explanation of:

 - How your graphics answers the question.
 - Explain the graphing choices you made for each (how does it follow OCAR?).

Provide code to the rest of the class reproducing your graphs. (Attach your RMD to the discussion post with the image.)