---
title: "Lecture 11 - Motion Notebook"
author: "Lindsay Waldrop"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

```{r setup}
# New libraries to install: 
library(av)
library(gganimate)

# Existing libraries to call:
library(ggplot2)
```

# Motion in R

There are many options for creating motion in R. For a basic plot, simple animation requires frames in succession where movements are broken up in to smaller bits. This code is an example that produces individual PNG images in a folder called anim_plots:

```{r}
dir.create("anim_plots", showWarnings = FALSE)

dist <- function(x1, x2, n, t) ((x2 - x1)/n)*t

plot_it <- function(dat, dat_old1, dat_old2){
   p <- ggplot(dat, aes(x, y)) + geom_point() + 
    geom_point(data = dat_old1, mapping = aes(x, y), alpha = 0.3) +
    geom_point(data = dat_old2, mapping = aes(x, y), alpha = 0.1) +
    xlim(0, 1) + ylim(0, 1) + xlab("X") + ylab("Y") + 
    theme_bw()
   return(p)
}

num_pts <- 100
n_frames <- 60*3

x1 <- runif(num_pts)
x2 <- runif(num_pts)
y1 <- runif(num_pts)
y2 <- runif(num_pts)
dat_old1 <- data.frame("x" = x1, "y" = y1)
dat_old2 <- data.frame("x" = x1, "y" = y1)
p <- list()
for (j in 1:n_frames){
  dat <- data.frame("x" = x1 + dist(x1, x2, n_frames, j), 
                    "y" = y1 + dist(y1, y2, n_frames, j))
  p[[j]] <- plot_it(dat, dat_old1, dat_old2)
  ggsave(filename = paste0("anim_plots/dotmove-", j, ".png"), p[[j]])
  dat_old2 <- dat_old1
  dat_old1 <- dat
}
  
```

Stitching these together gives a video of points traveling between two states (x1, y1) and (x2, y2). You can set both the number of frames and the number of points to plot.

## Using the `av` package to create a video

The `av` packages operates with the same functionality as the command-line tool `ffmpeg`. It works directly through R, which makes it useful for dealing with creating videos with R graphics.

Example of encoding video with the previously created PNG files:

```{r include=FALSE}
png_files <- sprintf("anim_plots/dotmove-%01d.png", 1:180)

av_encode_video(png_files, "av-anim1.mp4", framerate = 60)
```

It also has the functionality of being able to directly open devices that are used to save videos.

```{r include=FALSE}
av_capture_graphics(print(p), "av-anim2.mp4", 
                      1280, 720, res = 300, framerate = 30)
```

R Bloggers has a good little spot on the `av` package here: <https://www.r-bloggers.com/2018/10/the-av-package-production-quality-video-in-r/>

## Using the `gganimate` package

The tidyverse has its own animate package that works nicely with `ggplot2` which is `gganimate`. It can output gifs and videos and do automatic transitions between ggplot graph elements. More info is here: <https://gganimate.com/>

Examples from the vignette:

```{r}
p <- ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  # Here comes the gganimate code
  transition_states(
    gear,
    transition_length = 2,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('sine-in-out')
p
```

The `anim_save()` function can encode this into a variety of different things.

```{r}
anim_save("mtcars-anim1.gif", animation = p)
```

The `av` package also has a nice output renderer that works with `gganimate`:

```{r}
animate(p, renderer = av_renderer('mtcars-anim2.mp4'), 
    width = 1280, height = 720, res = 200, fps = 30)
```
