---
title: "Lecture 5 - Luminance & ggplot2 Notebook"
author: "Lindsay Waldrop"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

```{r setup}
#new libraries to install: 
library(lattice)
library(ggplot2)
library(tidyr)
library(tibble)
```
# `ggplot2` package

## Plotting Systems in R

### Base graphics and the `grid` package

Example of a plot in base graphics using the `mtcars` data set: 

```{r, include=TRUE}
plot(x = mtcars$mpg, y = mtcars$hp, 
     col = "black", bg = "red", pch = 21)
```

### The Trellis system in the `lattice` package

Example of a density plot in the `lattice` package:

```{r,include=TRUE}
densityplot(~mpg, data = mtcars,
            main = "Density Plot",
            xlab = "Miles per Gallon")
```
### The Grammar of Graphics `ggplot2` package

Example of a plot using `ggplot2`:

```{r, include=TRUE}
ggplot(mtcars, aes(x = mpg, y = hp, fill = factor(cyl))) + 
  geom_point(pch = 21, size = 2)  +
  xlab("Miles per Gallon") + ylab("Horsepower")
```


## Changing Attributes

### Changing Basic Plot Attributes

Here are some example plots in the lecture, simply comment or uncomment the ones you want to see plotted. 

Plot and axis labels: 

```{r, include=TRUE}
p <- ggplot(mtcars, aes(x = mpg, y = hp)) +
  geom_point()
p + xlab("Miles Per Gallon") # X-axis label
p + ylab("Horsepower") # Y-axis label
p + ggtitle("Horsepower vs MPG from MTCars") # chart title
```

Axis scales: 
```{r,include=TRUE}
p + xlim(0, 50) # altering limits of x-axis
p + ylim(0, 500) # altering limits of y-axis
# Note: no color to alter
```

Formatting plot area and text: 
```{r, include=TRUE}
# Alters size of text throughout plot and angel of the text to the x-axis.
p + theme(text = element_text(size = 20),
          axis.text.x = element_text(angle = 90, hjust = 1)) 
```

Minimal theme: 
```{r,include=TRUE}
p + theme_minimal()
```

### Changing Aesthetic Values

Changing point size, shape, color, and fill independently of mapping:
```{r,include=TRUE}
p2 <- ggplot(mtcars, aes(x = mpg, y = hp)) 
p2 + geom_point(size = 2)
p2 + geom_point(shape = 3)
p2 + geom_point(color = "red")
p2 + geom_point(pch = 21, color = "black", fill = "red")
```


### Grouping and Aesthetic Values

In order to have things like color, point, or size alter with the values of one of your data columns, you have to *map* this data onto a value. Here are three examples of how to map the `cyl` column from mtcars onto either fill, color, or shape. Uncomment the line you'd like to see.

```{r,include=TRUE}
# color:
ggplot(mtcars, aes(x = mpg, y = hp, color = factor(cyl))) + geom_point()
# fill:
ggplot(mtcars, aes(x = mpg, y = hp, fill = factor(cyl))) + geom_point(shape = 21)
# shape:
ggplot(mtcars, aes(x = mpg, y = hp, shape = factor(cyl))) + geom_point()
```

Mapping something like color in the initial `ggplot()` call of `aes()` will map that variable onto every geom that you add: 
```{r,include=TRUE}
ggplot(mtcars, aes(x = mpg, y = hp, color = factor(cyl))) + 
  geom_point() + geom_line()
```

If you would like an aesthetic value to change with one of your data columns, this MUST be done inside `aes()`. The following maps incorrectly and will not produce a legend:
```{r,include=TRUE}
ggplot(mtcars, aes(x = mpg, y = hp)) + geom_point(color = factor(mtcars$cyl))
```

#### `ggplot` exercise

Use the Orange data set to create a scatterplot with `ggplot2` of circumference versus tree age. Add to the plot: 
 1. Adding axis labels and a title.
 2. Changing the color, shape, and size of the points.
 3. Rotating the x-axis labels by 45 degrees.
 
```{r}
data("Orange")

```


## Data Shape and `ggplot`

To reap the most benefits from the `ggplot` package, many times it will require you to reshape your data from a traditional "wide" format to a "long" format. 

Reshape the `Loblolly` dataset into the wide format using `pivot_wider` from the `tidyr` package:
```{r,include=TRUE}
data("Loblolly")
Loblolly.wide <- pivot_wider(Loblolly, 
                             names_from = age, 
                             values_from = height)
Loblolly.wide
```
Now convert Loblolly into the long format, preferred by `ggplot`, using `pivot_longer` in the `tidyr` package:

```{r,include=TRUE}
Loblolly.long <- pivot_longer(Loblolly.wide, 
                              cols = !Seed, 
                              names_to = "age", 
                              values_to = "Height", 
                              names_transform = list(age = as.numeric))
Loblolly.long
```

It's not always necessary to reshape your data frame, it depends on what you want to plot and how you want it mapped!


### Plotting with wide vs long data

Using the two loblolly data frames, here are two plots. 

First, in the wide format, we can plot individual seeds at individual years, but are unable to plot something perhaps more useful, like height versus age. You MUST map each column to x and y, which limits your options in the wide format.
```{r,include=TRUE}
ggplot(Loblolly.wide, aes(x = Seed, y = `3`)) + geom_point()
```
However, in the long format useful things like height versus age are easily plotted out: 
```{r,include=TRUE}
ggplot(Loblolly.long, aes(x = factor(age), y = Height, color = Seed)) +
  geom_point(pch = 19) 
```

It's also possible to run other geoms like `geom_boxplot()`, which will calculate the means and quartiles for each age and plot them: 

```{r,include=TRUE}
ggplot(Loblolly.long, aes(x = factor(age), y = Height)) + 
  geom_boxplot()
```

Another popular plot is a violin plot, where the distribution of data is represented by the width of the violin: 

```{r,include=TRUE}
ggplot(Loblolly.long, aes(x = factor(age), y = Height)) + 
  geom_violin()
```

#### Long/wide exercise here

1. Reshape the `fish_encounters` data set (in the `tidyr` package) into a wide format.
```{r, include=TRUE}
data("fish_encounters", package = "tidyr")

```

2. Reshape the `relig_income` data set (in the `tidyr` package) into a long format.
```{r, include=TRUE}
data("relig_income", package = "tidyr")

```

## Adding Data to Existing Plots

One of the more clunky things about ggplot is that adding data from an outside data set is a bit awkward. You can do this, but you have to define the new data set you're using and the mapping within the geom. 

For this example, we will use the `Loblolly.wide` data frame and add another column to the graph. 
```{r,include=TRUE}
# Create and store graph with year 3 data:
p2 <- ggplot(Loblolly.wide, aes(x = Seed, y = `3`)) + 
  geom_point()
# Add year 5 data to the existing graph p2:
p2 + geom_point(data = Loblolly.wide, 
                mapping = aes(x = Seed, y = `5`),
                color = "red")
```
This is super clunky and will not give you an automatic legend. Try to avoid adding data in this manner unless it is unavoidable!

Sometimes this is necessary because of the nature of the data sets you'd like to integrate.


# Luminance

Code accompanying lecture on luminance.

### Color map exercise

Here, we'll use the volcano data set to plot a color map. 
```{r}
data(volcano)

volcano_long <- as.data.frame(volcano) %>% 
  rownames_to_column(var = "Y") %>% 
  pivot_longer(-Y, names_to = "X", values_to = "value", names_prefix = "V") %>%
  transform(Y = as.numeric(Y)) %>%
  transform(X = as.numeric(X))
  
ggplot(volcano_long, aes(X,Y, fill = value)) + geom_tile()

```



