---
title: "Lecture 21 - ggplot2 hacks"
author: "Lindsay Waldrop"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_notebook
---

```{r setup}
# New libraries to install: 
# will be included in each section/chunk

# Existing libraries to call:
library(ggplot2)
```

# ggplot2 and the tidyverse

`ggplot2` is a package for data visualization in the tidyverse, a suite of packages in R designed to aid in reading, analyzing, and visualizing data. More info on the tidyverse: https://www.tidyverse.org/ 

There are many package extensions of `ggplot2`, those that are considered compatible and using the same structure as other packages in the tidyverse. A convenient list of these can be found here, with links to the package information and vignettes: https://exts.ggplot2.tidyverse.org/gallery/

Here are some of my personal favorites! 

### patchwork

`patchwork` makes it really easy to combine separate ggplot objects into a single figure. The +, /, ( ) operators are used to combine each of these, with more fine-grain control with internal options. 

https://github.com/thomasp85/patchwork 

```{r}
#install.packages("patchwork")
library(patchwork)

data("diamonds")

p_carat <- ggplot(diamonds, aes(x=carat, y = price, color = color)) + 
  geom_point() + 
  xlab("Diamond carat") + ylab("Diamond Price (USD)") +
  theme_minimal()

p_cut <- ggplot(diamonds, aes(x = cut, y = price, color = color, fill = color)) + 
  geom_boxplot() +
  xlab("Diamond cut") + ylab("Diamond Price (USD)") +
  theme_minimal()

p_carat + p_cut
#ggsave("patchwork_sidebyside.png", last_plot())
 # or 

p_carat / p_cut
#ggsave("patchwork_topandbottom.png", last_plot())
```

## ggstatsplot

`ggstatsplot` has a suite of common statistical comparisons built in to a ggplot visualization. Statistical tests are automatically run and reported, along with statistical objects on the visualization.

https://github.com/IndrajeetPatil/ggstatsplot 

Example from the vignette: 
```{r}
#install.packages("ggstatsplot")
library(ggstatsplot)

set.seed(123)

ggbetweenstats(
  data  = iris,
  x     = Species,
  y     = Sepal.Length,
  title = "Distribution of sepal length across Iris species"
)
#ggsave("ggstatsplot-example.png", last_plot())
```

## ggrepel

`ggrepel` creates labels on ggplots that do not overlap with data or other labels. This seems sort of trivial, but it’s highly annoying when you have to deal with it! 

https://github.com/slowkow/ggrepel 

Example from the vignette: 
```{r}
#install.packages("ggrepel")
library(ggrepel)

ggplot(mtcars, aes(wt, mpg, label = rownames(mtcars))) +
  geom_text_repel() +
  geom_point(color = 'red') +
  theme_classic(base_size = 16)
ggsave("ggrepel-example.png", last_plot())
```

## ggraph

`ggraph` is designed to help plot relational and/or data structures in a ggplot type format. This is a great package if you want to plot network diagrams, an otherwise difficult task in ggplot. 

https://github.com/thomasp85/ggraph 

```{r}
#install.packages(c("ggraph", "tidygraph"))
library(ggraph)
library(tidygraph)

# Create graph of highschool friendships
graph <- as_tbl_graph(highschool) %>% 
    mutate(Popularity = centrality_degree(mode = 'in'))

# plot using ggraph
ggraph(graph, layout = 'kk') + 
    geom_edge_fan(aes(alpha = after_stat(index)), show.legend = FALSE) + 
    geom_node_point(aes(size = Popularity)) + 
    facet_edges(~year) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')
#ggsave("ggraph-example.png", last_plot())
```

## ggalluvial

ggalluvial creates alluvial plots for visualizing proportions or frequency distributions. 

https://github.com/corybrunson/ggalluvial 

Example from https://cran.r-project.org/web/packages/ggalluvial/vignettes/ggalluvial.html: 
```{r}
#install.packages("ggalluvial")
library(ggalluvial)
majors$curriculum <-as.factor(majors$curriculum)
ggplot(majors,
       aes(x = semester, stratum = curriculum, alluvium = student,
           fill = curriculum, label = curriculum)) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  geom_flow(stat = "alluvium", lode.guidance = "frontback",
            color = "darkgray") +
  geom_stratum() +
  theme(legend.position = "bottom") +
  ggtitle("student curricula across several semesters")
#ggsave("alluvial-example.png", last_plot())
```

## ggspectra

`ggspectra` creates spectra plots and statistics for light wave data. 

https://docs.r4photobiology.info/ggspectra/ 

```{r}
#install.packages("ggspectra")
library(ggspectra)

ggplot(sun.spct, unit.out = "photon") +
  geom_spct() +
  scale_y_s.q.irrad_continuous() +
  scale_x_wl_continuous(sec.axis = sec_axis_w_number()) +
  theme_bw()
ggsave("ggspectra-example.png", last_plot())
```